
---
title: "WorldBankCommodities"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, fig.height=14, fig.width=14, include=FALSE, messsage=FALSE}
library(xts)
# setwd("~/Google Drive/CommodityMeanReversion")
# data in in a gsheet, must output it to csv
#
timeLag <- 120

# read in the data from the csv file
x <- read.csv('WorldBankData.csv')

# replace '..' and '0.00' with NA
x[x=='..'] <- NA
x[x=='0.00'] <- NA

# create a date vector
#dateDum <- as.Date(x[,1])
dateDum=as.Date(x[,1], '%m/%d/%Y')
# take out the date column
x <- x[ ,-1 ]

# calculate the deflator
deflator <- x$usd / x$PCEPI * tail(x$PCEPI,1) / tail(x$usd,1)
```

# ```{r mm, echo=FALSE, fig.height=8, fig.width=9, warning=FALSE, include=FALSE, results='asis'}
``` {r echo=FALSE, warning=FALSE, fig.height=8, fig.width=9, results='asis'}
# getting the plots

for ( i in 1:77) {
    if (is.na(tail(x[,i],1))) next
  series <- as.numeric( as.character(x[,i]) ) 
  com <- series * deflator
  XTScom <- xts(com,dateDum)
  XTScom <- na.omit(XTScom)
  blue = FALSE
  if ( index(head(XTScom,1)) < '1972-01-01' ) {
    blue = TRUE
    blueLen <- length(XTScom['/1972-01-01'])
}
com <- as.numeric(XTScom)

  du1 <- com[-(1:timeLag)]
  if ( timeLag >= length(com) ) next
  du2 <- com[1:(length(com)-timeLag)]
  ret <- du1 / du2
  title <- names(x[i])
  xTitle <- 'Price in 2017 $'
  xTitle <- paste('Price in', format(Sys.Date(), '%Y'), '$')
  xAxis  <- c(min(com), max(com))
  yAxis  <- c(min(ret), max(ret))
  yTitle <- paste( timeLag/12, 'Year Forward Return', sep='' )
  g=length(ret)
  if (g>320) br = g-320 else br=5
  if (blue == TRUE) {
    plot(ret[blueLen:br]~du2[blueLen:br], pch=1, main=title, log='xy'
         , xlab=xTitle, ylab=yTitle, xlim=xAxis, ylim=yAxis)
    points(ret[br:g] ~ du2[br:g], col='red')
    points(ret[1:blueLen] ~ du2[1:blueLen], col='blue')
}    
  if (blue == FALSE) {
    plot(ret[1:br]~du2[1:br], pch=1, main=title, log='xy'
         , xlab=xTitle, ylab=yTitle, xlim=xAxis, ylim=yAxis)
    points(ret[br:g] ~ du2[br:g], col='red')
}    
    
  abline( v = tail(com,1) , col='red',lwd=2 )
  grid(col='blue', lwd = 1, lty = 3)
  avg <- mean(com) ; std <- sd(com)
  numSD <- format( ( tail(com,1) - avg ) / std, digits = 3 )
  ddum <- 'over'
  if (numSD < 0.0) ddum <- 'under'
  print( paste( title, 'is', numSD, 'stnd devs', ddum, 'the average   '))
  
  regY  <- log(ret)
  regX1 <- log(du2)
  regX2 <-  seq(1:length(ret))
  regX3 <- regX2 * regX2
  regX4 <- regX3 * regX2
  OLSregression <- lm( regY ~ regX1 ) #+ regX2 + regX3 + regX4 )
   rr=data.frame(regX1=log(tail(series,1)) ) #, regX2=log(dim(x)[1]), regX3=log(dim(x)[1]^2)
       #          , regX4=log(dim(x)[1]^3) )
    currForecast <- predict(OLSregression,newdata = rr)
    currForecast <- format( exp(currForecast), digits = 3)
    print(paste('Current 10-year forecast =', currForecast))
}

```
